---
title: "NY Housing Cleaning"
output: pdf_document
author: "Alex Stern"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Evaluation of Missingness
```{r message=FALSE}
setwd("~/Desktop/ds4001")
train <- read.csv("sales_manhattan_train_set.csv", stringsAsFactors = FALSE)
# ensure the same data manipulation is performed on the test set, seperately 
test <- read.csv("sales_manhattan_test_set.csv", stringsAsFactors = FALSE)
summary(train)
summary(train$EASE.MENT)
nrow(train)

# Tax class at present has missing values as well
# 2 is by far the mode for this class, therefore 2 will be imputed for all ""
train$TAX.CLASS.AT.PRESENT[which((train$TAX.CLASS.AT.PRESENT) == "")] <- "2"
test$TAX.CLASS.AT.PRESENT[which((test$TAX.CLASS.AT.PRESENT) == "")] <- "2"
# replace missing building class at present values with the mode
train$BUILDING.CLASS.AT.PRESENT[which((train$BUILDING.CLASS.AT.PRESENT) == "")] <- "R4"
test$BUILDING.CLASS.AT.PRESENT[which((test$BUILDING.CLASS.AT.PRESENT) == "")] <- "R4"
# TAX CLASS AT SALE should be a factor as well
train$TAX.CLASS.AT.TIME.OF.SALE <- as.factor(train$TAX.CLASS.AT.TIME.OF.SALE)
test$TAX.CLASS.AT.TIME.OF.SALE <- as.factor(test$TAX.CLASS.AT.TIME.OF.SALE)
# zip code should be a factor
train$ZIP.CODE <- as.factor(train$ZIP.CODE)
test$ZIP.CODE <- as.factor(test$ZIP.CODE)
# ensure 0 isn't used in calculating the mean
train$YEAR.BUILT[train$YEAR.BUILT == 0] <- NA
test$YEAR.BUILT[test$YEAR.BUILT == 0] <- NA
# replace missing year values with median (since a few buildings will be very old)
train$YEAR.BUILT[is.na(train$YEAR.BUILT)] <- median(train$YEAR.BUILT, na.rm = TRUE)
test$YEAR.BUILT[is.na(test$YEAR.BUILT)] <- median(test$YEAR.BUILT, na.rm = TRUE)
# Since the value of "EASE.MENT" in every row is NA, the feature can be dropped without losing any information, apt number should be dopped too due to lack of info it provides
# borough is the same for every row and can be dropped
drops <- c("EASE.MENT", "APARTMENT.NUMBER", "BOROUGH")
train <- train[ , !(names(train) %in% drops)]
test <- test[ , !(names(test) %in% drops)]
write.csv(unique(train$NEIGHBORHOOD), "neighborhoods.csv")
```


## Feature Manipulation and Creation
```{r}
library(lubridate)
library(ggplot2)
library(tidyverse)
train$sale_year <- year(mdy(train$SALE.DATE))
train$sale_month <- month(mdy(train$SALE.DATE))
train$sale_day <- day(mdy(train$SALE.DATE))

test$sale_year <- year(mdy(test$SALE.DATE))
test$sale_month <- month(mdy(test$SALE.DATE))
test$sale_day <- day(mdy(test$SALE.DATE))

train$sale_minus_built <- (train$sale_year - train$YEAR.BUILT)
test$sale_minus_built <- (test$sale_year - test$YEAR.BUILT)
# pull the year, month, and day as individual features
# drop origianl date column which no longer provides any value
# drop zip code (many missing values), address (little info)
# the vast majority have no land square feet
# half are missing gross square feet
# "LAND.SQUARE.FEET", "GROSS.SQUARE.FEET"
drops <- c("SALE.DATE", "ZIP.CODE", "ADDRESS", "BUILDING.CLASS.AT.PRESENT", "LOT", "TAX.CLASS.AT.PRESENT")
train <- train[ , !(names(train) %in% drops)]
test <- test[ , !(names(test) %in% drops)]

building_class_same <- c()
for (i in 1:nrow(train)) {
  if (train[i,"BUILDING.CLASS.CATEGORY"] == train[i,"BUILDING.CLASS.AT.TIME.OF.SALE"]) {
    building_class_same <- c(building_class_same, 1)
  }
  else {
    building_class_same <- c(building_class_same, 0)
  }
}
train$building_class_same <- building_class_same

building_class_same <- c()
for (i in 1:nrow(test)) {
  if (test[i,"BUILDING.CLASS.CATEGORY"] == test[i,"BUILDING.CLASS.AT.TIME.OF.SALE"]) {
    building_class_same <- c(building_class_same, 1)
  }
  else {
    building_class_same <- c(building_class_same, 0)
  }
}
test$building_class_same <- building_class_same

train <- transform(train, GROSS.SQUARE.FEET = as.numeric(GROSS.SQUARE.FEET))
train <- transform(train, LAND.SQUARE.FEET = as.numeric(LAND.SQUARE.FEET))
train<-train%>%mutate(
  GROSS.SQUARE.FEET=ifelse(GROSS.SQUARE.FEET<2,NA,GROSS.SQUARE.FEET),
  LAND.SQUARE.FEET=ifelse(LAND.SQUARE.FEET<2,NA,LAND.SQUARE.FEET)
)
train$GROSS.SQUARE.FEET[is.na(train$GROSS.SQUARE.FEET)] <- median(train$GROSS.SQUARE.FEET, na.rm = TRUE)
train$LAND.SQUARE.FEET[is.na(train$LAND.SQUARE.FEET)] <- median(train$LAND.SQUARE.FEET, na.rm = TRUE)

test <- transform(test, GROSS.SQUARE.FEET = as.numeric(GROSS.SQUARE.FEET))
test <- transform(test, LAND.SQUARE.FEET = as.numeric(LAND.SQUARE.FEET))
test<-test%>%mutate(
  GROSS.SQUARE.FEET=ifelse(GROSS.SQUARE.FEET<2,NA,GROSS.SQUARE.FEET),
  LAND.SQUARE.FEET=ifelse(LAND.SQUARE.FEET<2,NA,LAND.SQUARE.FEET)
)
test$GROSS.SQUARE.FEET[is.na(test$GROSS.SQUARE.FEET)] <- median(test$GROSS.SQUARE.FEET, na.rm = TRUE)
test$LAND.SQUARE.FEET[is.na(test$LAND.SQUARE.FEET)] <- median(test$LAND.SQUARE.FEET, na.rm = TRUE)

# from Shaun's data manipulation
train$high.end<-as.factor(train$high.end)
train$TAX.CLASS.AT.TIME.OF.SALE<-as.factor(train$TAX.CLASS.AT.TIME.OF.SALE)

train<-train%>%mutate(
  COMMERCIAL.UNITS=ifelse(COMMERCIAL.UNITS==0,F,T),
  RESIDENTIAL.UNITS=ifelse(RESIDENTIAL.UNITS==0,F,T),
  TOTAL.UNITS=ifelse(TOTAL.UNITS==0,F,T)
)

test$TAX.CLASS.AT.TIME.OF.SALE<-as.factor(test$TAX.CLASS.AT.TIME.OF.SALE)

test<-test%>%mutate(
  COMMERCIAL.UNITS=ifelse(COMMERCIAL.UNITS==0,F,T),
  RESIDENTIAL.UNITS=ifelse(RESIDENTIAL.UNITS==0,F,T),
  TOTAL.UNITS=ifelse(TOTAL.UNITS==0,F,T)
)
summary(train)
write_csv(train, "train_clean2.csv")
write_csv(test, "test_clean2.csv")
```
The following variables will be encoded into dummy categorical variables via a pipeline in Python before the final model is fit: NEIGHBORHOOD, BUILDING.CLASS.CATEGORY, TAX.CLASS.AT.PRESENT, BUILDING.CLASS.AT.PRESENT, TAX.CLASS.AT.TIME.OF.SALE, BUILDING.CLASS.AT.TIME.OF.SALE, RESIDENTIAL.UNITS, COMMERCIAL.UNITS, TOTAL.UNITS, building_class_same (is already dummy coded)






















