---
title: "Project 2 Final"
author: "Siyeon Shaun Kim (sk5ps), Alex Stern, Josh Eilen"
date: "11/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

##Data Cleaning
```{r}
train<-read.csv("sales_manhattan_train_set.csv")
test<-read.csv("sales_manhattan_test_set.csv")

train$high.end<-as.factor(train$high.end)
train$TAX.CLASS.AT.TIME.OF.SALE<-as.factor(train$TAX.CLASS.AT.TIME.OF.SALE)
train$LAND.SQUARE.FEET<-as.numeric(train$LAND.SQUARE.FEET)
train$GROSS.SQUARE.FEET<-as.numeric(train$GROSS.SQUARE.FEET)

train<-train%>%mutate(
  COMMERCIAL.UNITS=ifelse(COMMERCIAL.UNITS==0,F,T),
  RESIDENTIAL.UNITS=ifelse(RESIDENTIAL.UNITS==0,F,T),
  TOTAL.UNITS=ifelse(TOTAL.UNITS==0,F,T),
  YEAR.BUILT=ifelse(YEAR.BUILT==0,1932,YEAR.BUILT)
  
)

train_prepared<-train%>%select(
  -BOROUGH, -EASE.MENT, -ADDRESS
)


summary(train_prepared)
```
We got rid of BOROUGH, EASE.MENT, ADDRESS because BOROUGH variable has unanimous values, all the EASE.MENT values are NA, and ADDRESS will not be our main predictor. We also factorized high.end, TAX.CLASS.AT.TIME.OF.SALE, RESIDENTIAL UNITS,and COMMERCIAL UNITS. Now, we implemented the same data process to the test training set. 

```{r}
test$TAX.CLASS.AT.TIME.OF.SALE<-as.factor(test$TAX.CLASS.AT.TIME.OF.SALE)
test$LAND.SQUARE.FEET<-as.numeric(test$LAND.SQUARE.FEET)
test$GROSS.SQUARE.FEET<-as.numeric(test$GROSS.SQUARE.FEET)

test<-test%>%mutate(
  COMMERCIAL.UNITS=ifelse(COMMERCIAL.UNITS==0,F,T),
  RESIDENTIAL.UNITS=ifelse(RESIDENTIAL.UNITS==0,F,T),
  TOTAL.UNITS=ifelse(TOTAL.UNITS==0,F,T)
)

test_prepared<-test%>%select(
  -BOROUGH, -EASE.MENT, -ADDRESS,-ZIP.CODE
)

summary(test_prepared)
```

## Exploratory Data Analysis 
```{r}
ggplot(data=train_prepared)+geom_point(mapping =aes(x=BLOCK, 
                                                       y=GROSS.SQUARE.FEET,color=high.end))
ggplot(data=train_prepared)+geom_point(mapping =aes(x=YEAR.BUILT, 
                                                    y=BLOCK,color=high.end))
ggplot(data=train_prepared)+geom_point(mapping =aes(x=YEAR.BUILT, 
                                                    y=GROSS.SQUARE.FEET,color=high.end))
ggplot(data=train_prepared%>%filter(NEIGHBORHOOD=="MIDTOWN EAST"))+
  geom_point(mapping =aes(x=LAND.SQUARE.FEET,y=GROSS.SQUARE.FEET,color=high.end))
```
This will be pretty difficult process building models because there is no apparent correlation between different features and its relationship with the high.end feature. 


## Logistic Regression
```{r}
library(glmnet)
library(ROCR)

summary(train_prepared)
summary(test_prepared)
attach(train_prepared)

model<-glm(high.end~ NEIGHBORHOOD+BUILDING.CLASS.CATEGORY+TAX.CLASS.AT.PRESENT+
             BLOCK+LOT+RESIDENTIAL.UNITS+
             COMMERCIAL.UNITS+LAND.SQUARE.FEET+GROSS.SQUARE.FEET+YEAR.BUILT
             +TAX.CLASS.AT.TIME.OF.SALE,family=binomial)
summary(model)
pred<-predict(model,newdata=train_prepared)

rates<-prediction(pred, train_prepared$high.end)
roc_result<-performance(rates,measure="tpr", x.measure="fpr")
plot(roc_result, main="ROC Curve")+lines(x = c(0,1), y = c(0,1), col="red")

confusion.mat<-table(train_prepared$high.end,pred > 0.5)
confusion.mat

(8174+2021)/(14150)


## model 2 using Logistic regression
model2<-glm(high.end~ NEIGHBORHOOD+BUILDING.CLASS.CATEGORY+
             BLOCK+LOT+BUILDING.CLASS.AT.PRESENT+RESIDENTIAL.UNITS+
             COMMERCIAL.UNITS+GROSS.SQUARE.FEET+
             TAX.CLASS.AT.TIME.OF.SALE+BUILDING.CLASS.AT.TIME.OF.SALE,family=binomial)

pred2<-predict(model,newdata=train_prepared)

rates2<-prediction(pred2, train_prepared$high.end)
roc_result2<-performance(rates2,measure="tpr", x.measure="fpr")
plot(roc_result2, main="ROC Curve")+lines(x = c(0,1), y = c(0,1), col="red")

confusion.mat2<-table(train_prepared$high.end,pred2 > 0.5)
confusion.mat2
```


## Support Vector Machine
```{r}
source("datacleaning.r")
library(boot) ##for cv.glm function
library(MASS) ##for lda function
library(ipred) ##for errorest function
library(klaR) ##for partimat

lda.data <- lda(high.end~BLOCK+LOT+LAND.SQUARE.FEET+
                  YEAR.BUILT+GROSS.SQUARE.FEET
                ,train_prepared)
test_pred<-predict(lda.data,newdata=train_prepared)
table(train$high.end,test_pred$class)
mean(train$high.end==test_pred$class)
plot(lda.data, col = as.integer(train_prepared$high.end), main="LDA")

pred<-predict(lda.data,newdata=test_prepared)
high.end<-pred$class


#### qda

qda.data <- qda(high.end~BLOCK+LOT+LAND.SQUARE.FEET+
                  YEAR.BUILT+GROSS.SQUARE.FEET
                ,train_prepared)
test_pred<-predict(qda.data,newdata=train_prepared)
table(train$high.end,test_pred$class)
mean(train$high.end==test_pred$class)


pred<-predict(lda.data,newdata=test_prepared)
high.end<-pred$class
```
Support Vector Machine only takes quantitative features, so we knew that the model prediction will be limited. 

##Random Forest
```{r}
library(tree)
library(randomForest) ##for random forests (and bagging)
library(gbm) ##for boosting


rnd.class<-randomForest(high.end~ NEIGHBORHOOD +BUILDING.CLASS.CATEGORY +
                         TAX.CLASS.AT.PRESENT+BLOCK + LOT + RESIDENTIAL.UNITS 
                        + COMMERCIAL.UNITS + LAND.SQUARE.FEET+
                          TAX.CLASS.AT.TIME.OF.SALE+
                          GROSS.SQUARE.FEET,data=train_prepared,mtry=6,
                        importance=TRUE) 
rnd.class
importance(rnd.class)

levels(test_prepared$NEIGHBORHOOD) <- levels(train_prepared$NEIGHBORHOOD)
levels(test_prepared$BUILDING.CLASS.CATEGORY) <- levels(train_prepared$BUILDING.CLASS.CATEGORY)
levels(test_prepared$BLOCK) <- levels(train_prepared$BLOCK)
levels(test_prepared$LOT) <- levels(train_prepared$LOT)
levels(test_prepared$RESIDENTIAL.UNITS) <- levels(train_prepared$RESIDENTIAL.UNITS)
levels(test_prepared$COMMERCIAL.UNITS) <- levels(train_prepared$COMMERCIAL.UNITS)
levels(test_prepared$GROSS.SQUARE.FEET) <- levels(train_prepared$GROSS.SQUARE.FEET)
levels(test_prepared$TAX.CLASS.AT.PRESENT) <- levels(train_prepared$TAX.CLASS.AT.PRESENT)
levels(test_prepared$TAX.CLASS.AT.TIME.OF.SALE) <- levels(train_prepared$TAX.CLASS.AT.TIME.OF.SALE)
levels(test_prepared$LAND.SQUARE.FEET) <- levels(train_prepared$LAND.SQUARE.FEET)

rnd.pred<-predict(rnd.class,newdata=test_prepared) 
```
